% §11 LLM Usage Statement — outside page-limit per ACSAC CFP
% The statement must cover BOTH roles: LLM as subject of study AND LLM as writing aid.

\section*{LLM Usage Statement}
\label{sec:llm-usage}

This paper involves Large Language Models in two distinct roles.

\paragraph{LLM as subject of study}
The experimental work evaluates the penetration testing capabilities of LLM
agents using \texttt{MiniMax-M2.7}~\cite{minimax27} as the shared model across
\harnessname\ and all baselines (CAI, VulnBot, PentestGPT) to isolate the
architectural contribution from the model variable (see \S\ref{sec:setup}).
All inference calls were made through MiniMax's official API under standard
terms of service. No model weights were fine-tuned; all systems were used
out-of-the-box with identical tool schemas and prompt templates per system.
Prompt templates, full conversation transcripts, and per-phase token counts
are released as part of the artifact.

\paragraph{LLM as writing aid}
During manuscript preparation, the authors used current frontier LLMs
(Anthropic Claude Opus 4.7, OpenAI GPT-5.5) for: (i)~drafting candidate
section outlines subsequently edited by the authors; (ii)~copy-editing for
grammar, clarity, and concision; (iii)~consistency checks on terminology and
notation. No section was written end-to-end by an LLM. All claims,
experimental results, related-work positioning, and scientific interpretations
are the authors' own. Every LLM-suggested citation was verified by retrieving
the cited work.
