% IoTChainBench / LANCE — ACSAC 2026 submission
% Target format: 11 pages double-column IEEEtran conference compsoc
% Double-blind review — NO author info before camera-ready.
% Single-file version (sections inlined from sections/*.tex)

\documentclass[conference,compsoc]{IEEEtran}

% ── packages ──────────────────────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\renewcommand{\ttdefault}{cmtt}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath,amssymb}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit}
\usepackage{xspace}
\usepackage{verbatim}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}
\usepackage[hidelinks,breaklinks=true]{hyperref}
\usetikzlibrary{arrows.meta,fit,backgrounds}

% ── macros ────────────────────────────────────────────────────────────────
\newcommand{\systemname}{IoTChainBench\xspace}
\newcommand{\harnessname}{\textsc{LANCE}\xspace}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}
\newcommand{\note}[1]{\textcolor{blue}{\textit{[note: #1]}}}

\providecommand{\Cref}[1]{Section~\ref{#1}}
\providecommand{\cref}[1]{\S\ref{#1}}

\newenvironment{yamlquote}{\par\vspace{.3em}\footnotesize\verbatim}{\endverbatim\normalsize\par\vspace{.3em}}

% ── title & anonymous authors (ACSAC double-blind) ────────────────────────
\title{LANCE, a Network-Native LLM Agent and Benchmark for Multi-Hop IoT Penetration Testing}

% ── DOUBLE-BLIND: anonymized for submission ──────────────────────────────
% Camera-ready author block (uncomment for final version):
% \author{%
% \IEEEauthorblockN{Tanguy Vansnick}
% \IEEEauthorblockA{\textit{Faculty of Engineering - ILIA} \\
% \textit{University of Mons}\\
% Rue de Houdain 9, 7000 Mons, Belgium \\
% tanguy.vansnick@umons.ac.be}
% \and
% \IEEEauthorblockN{Gaspard Menou}
% \IEEEauthorblockA{\textit{Centrale Marseille}\\
% Marseille, France}
% \and
% \IEEEauthorblockN{Mathis Delehouzée}
% \IEEEauthorblockA{\textit{Faculty of Engineering - Polytech} \\
% \textit{University of Mons}\\
% Mons, Belgium}
% \and
% \IEEEauthorblockN{Saïd Mahmoudi}
% \IEEEauthorblockA{\textit{Faculty of Engineering - Polytech} \\
% \textit{University of Mons}\\
% Mons, Belgium}
% }
\author{\IEEEauthorblockN{Anonymous Authors}
\IEEEauthorblockA{Paper~\#XXXX, Anonymized for Double-Blind Review}}

\begin{document}
\maketitle

\begin{abstract}
Autonomous LLM penetration-testing agents are typically evaluated on
single hosts, yet IoT deployments expose risk through network reachability:
gateways, brokers, cameras, and OT services form attack paths whose security
depends on pivots across devices. We present \systemname, a reproducible
network-scale IoT benchmark with 15 Proxmox/Ansible scenarios, 145 devices
in total (133 in vulnerability-injected scenarios), and 229 injected
vulnerabilities mapped to OWASP IoT Top 10 and MITRE ATT\&CK for ICS.
We also introduce \harnessname, a six-phase network-native
harness that combines topology priors, active discovery, per-device triage,
per-vulnerability verification, intrusion attempts, and report generation.
The benchmark adds two metrics absent from single-host evaluations:
evidence-level coverage (L1 detection, L2 exploitation, L3 exfiltration) and
Multi-Hop Reach (MHR$_k$), which measures recall on vulnerabilities that
require at least $k$ pivots or zone transitions. We describe the benchmark,
evaluator, and experimental protocol for isolating architectural effects
independently from model choice. On the 10 main scenarios, \harnessname{}
achieves 0.962 recall, 0.922 precision, 0.940 F1, and 86.1\%
CVSS-weighted score. On external single-host corpora, a blind AutoPenBench
run captures 7/33 flags (21.2\%), while an informed run captures 9/33
(27.3\%); a full 191-case Vulhub run gives a conservative 39--48\%
useful-finding lower bound.
\end{abstract}

\begin{IEEEkeywords}
IoT security, penetration testing, large language models, autonomous agents,
benchmark, cyber-physical systems
\end{IEEEkeywords}

% ══════════════════════════════════════════════════════════════════════════
% §1 Introduction
% ══════════════════════════════════════════════════════════════════════════
\section{Introduction}
\label{sec:intro}

IoT deployments are compromised through paths, not just devices.
Mirai~\cite{antonakakis17}, Verkada~\cite{verkada21}, and
Stuxnet~\cite{langner11} each demonstrate this: impact came from
chained access across cameras, brokers, PLCs, and segmented boundaries.

Yet LLM penetration-testing agents are evaluated on a single host at a
time: isolated containers, VMs, or CTF services scored independently.
This gap is methodological, not merely architectural. Single-host corpora
cannot measure three capabilities central to IoT/OT security: finding
vulnerabilities behind pivots, reasoning over domain protocols (MQTT,
Modbus, CoAP, LoRaWAN), and distinguishing banner-level detection from
confirmed exploitation or exfiltration. Without a network-scale benchmark,
strong single-host scores overstate readiness for real deployments.

\medskip\noindent\textbf{Contributions.} This paper makes four contributions.
\begin{itemize}
  \item \textbf{IoTChainBench}: a reproducible network-scale benchmark with
  15 IoT/OT scenarios, 145 devices, and 229 injected vulnerabilities mapped
  to OWASP IoT Top~10 and MITRE ATT\&CK for ICS, deployed via Ansible with
  per-vulnerability ground truth.

  \item \textbf{LANCE}: a six-phase network-native harness
  (topology $\to$ discovery $\to$ triage $\to$ verification $\to$
  intrusion $\to$ report) achieving 0.940 F1 and 86.1\% CVSS-weighted
  score on the 10 main scenarios.

  \item \textbf{Evidence-aware metrics}: L1/L2/L3 scoring (detection,
  exploitation, exfiltration) and MHR$_k$ (recall as a function of
  firewall-hop depth), both absent from existing flag-capture benchmarks.

  \item \textbf{Controlled evaluation}: architectural comparison against
  adapted LLM-agent baselines under a fixed model and budget, with
  transfer validation on AutoPenBench and Vulhub.
\end{itemize}
All artifacts are released at \texttt{ANONYMIZED\_URL}.

% ══════════════════════════════════════════════════════════════════════════
% §2 Background and Related Work
% ══════════════════════════════════════════════════════════════════════════
\section{Related Work}
\label{sec:related}

This section reviews prior work relevant to our approach. We first survey LLM-driven penetration-testing agents and the benchmarks used to evaluate them, then discuss the classical attack-graph literature that informs \harnessname's topology-aware design, and finally review IoT firmware analysis and the security frameworks underlying our scenario taxonomy.

\subsection{LLM Penetration-Testing Agents}
\label{sec:related:llm}

\medskip\noindent\textbf{Single-agent systems.} Happe and Cito~\cite{happe23} were among the first to show that a single LLM driving a shell can autonomously perform privilege escalation on Linux VMs. \textsc{PentestGPT}~\cite{pentestgpt24} improved this with a three-module design over a \emph{Pentesting Task Tree}, outperforming direct GPT-3.5 and GPT-4 usage on HackTheBox and VulnHub targets. \textsc{HackingBuddyGPT}~\cite{hackingbuddygpt}, \textsc{AutoAttacker}~\cite{autoattacker24}, and \textsc{Nebula}~\cite{nebula24} extend this line as open-source tools.

\medskip\noindent\textbf{Multi-agent pipelines.} \textsc{VulnBot}~\cite{vulnbot25} decomposes pentesting into reconnaissance, scanning, and exploitation over a \emph{Penetration Task Graph}, gaining 21 points end-to-end over its single-agent baseline ($30.3\%$ on AutoPenBench with Llama~3.1-405B). \textsc{AutoPentester}~\cite{autopentester25} adds RAG across five agents for a $+27\%$ subtask gain over \textsc{PentestGPT} on HackTheBox. \textsc{PentestAgent}~\cite{pentestagent25} and \textsc{CAI}~\cite{cai25} generalise this blueprint with modular tool use and a public leaderboard.

\medskip\noindent\textbf{Model-level improvements.} \textsc{Pentest-R1}~\cite{pentestr125} applies two-stage reinforcement learning on reasoning traces. \textsc{xOffense}~\cite{xoffense25} fine-tunes Qwen3-32B on Chain-of-Thought penetration testing data, reaching $79.17\%$ subtask completion on AutoPenBench. \textsc{CheckMate}~\cite{checkmate25} couples an LLM with a \emph{Classical Planning+} module, gaining more than $20\%$ on Vulhub over a Claude Code + Sonnet~4.5 baseline.

\medskip\noindent\textbf{Common limitation.} Despite this diversity, reported
evaluations remain single-target or challenge-level. They do not score subnet
reachability, pivot depth, or per-vulnerability progress across IoT/OT
protocols (MQTT, Modbus, LoRaWAN, CoAP), which is the gap
\textsc{IoTChainBench} and \textsc{LANCE} address.

\subsection{LLM Cybersecurity Benchmarks}
\label{sec:related:bench}

\medskip\noindent\textbf{Challenge-level corpora.} \textsc{AutoPenBench}~\cite{autopenbench24}
is the dominant evaluation substrate: 33 Docker-container tasks with milestone
and flag-based scoring, used by \textsc{VulnBot}, \textsc{Pentest-R1}, and
\textsc{xOffense}. \textsc{Vulhub}~\cite{vulhub} provides reproducible
vulnerable environments used by \textsc{CheckMate}. \textsc{CAIBench}~\cite{caibench25},
\textsc{NYU CTF Bench}~\cite{nyuctfbench24}, and \textsc{Cybench}~\cite{cybench24}
extend coverage to CTFs, attack-and-defense ranges, and knowledge tests. These
corpora score at challenge or flag level. They do not provide per-vulnerability
ground truth tied to reachability, hop depth, or IoT/OT protocols.

\medskip\noindent\textbf{Multi-host labs.} \textsc{Ludus}~\cite{ludus} and
\textsc{GOAD}~\cite{goad} deliver Proxmox+Ansible deployments of enterprise
Active Directory networks. They are labs and training ranges, not scored LLM
benchmarks. Both model multi-host compromise and lateral movement, but neither
ships a per-vulnerability ground-truth schema, hop-depth annotations, or an
evaluator. Both also target enterprise AD rather than IoT/OT protocols. To our
knowledge, no prior benchmark combines multi-host reachability, IoT/OT protocol
coverage, per-vulnerability ground truth, and evidence-level scoring beyond
binary flag capture (Table~\ref{tab:related}).

\subsection{Attack Graphs and Network Reachability}
\label{sec:related:graph}

Representing a network as a graph for security analysis predates deep learning
by decades. Phillips and Swiler~\cite{phillips98} introduced graph-based
network-vulnerability analysis. Sheyner~\emph{et al.}~\cite{sheyner02}
developed automated attack-graph generation via model checking, and
\textsc{MulVAL}~\cite{mulval05} cast the problem as Datalog inference. These
systems compute reachable attacker states from a \emph{static} encoding of
host configurations and exposed services. \textsc{LANCE} borrows this
reachability abstraction but operates on a \emph{live} network: edges are
validated by probes, the LLM proposes the next action, and the graph is updated
as hosts are discovered. Classical attack-graph generation is therefore
complementary rather than competing.

\subsection{IoT Security Context}
\label{sec:related:iot}

Scenario design and ground-truth taxonomy follow three references: OWASP IoT Top~10~\cite{owasp-iot} for vulnerability categories, MITRE ATT\&CK for ICS~\cite{mitre-ics} for OT-tactic annotations, and NIST SP\,800-82~Rev.\,3~\cite{nist80082} for OT security and segmentation conventions.

Empirical IoT-security artefacts fall into two families. \textbf{Network-traffic
datasets} (IoT-23~\cite{iot23}, TON\_IoT~\cite{toniot20},
Edge-IIoTset~\cite{edgeiiotset22}) provide labeled traces for intrusion
detection. They are not deployable pentest targets with per-vulnerability
ground truth or controllable reachability. \textbf{Firmware-focused work}
(Costin~\emph{et al.}~\cite{costin14}, \textsc{Firmadyne}~\cite{firmadyne16},
\textsc{FirmAE}~\cite{firmae20}, \textsc{IoTGoat}~\cite{iotgoat}) exercises
single-device firmware analysis. These datasets measure detection or
firmware-analysis capability. \textsc{IoTChainBench} instead measures active
security assessment on deployed multi-device networks with firewall-enforced
reachability.

\begin{table}[h!]
\centering
\caption{Gap analysis. \emph{Scale}: single host vs.\ multi-device network.
\emph{IoT/OT}: application-layer IoT/OT protocols on the test surface.
\emph{GT}: ground-truth granularity (flag, task-step, per-vulnerability, none).
\emph{MH}: multi-hop scoring.}
\label{tab:related}
\setlength{\tabcolsep}{4pt}
\scriptsize
\begin{tabular}{@{}lcccc@{}}
\toprule
System & Scale & IoT/OT & GT & MH \\
\midrule
\multicolumn{5}{@{}l}{\emph{LLM pentest agents}} \\
\midrule
Happe \& Cito~\cite{happe23}         & 1 host  & \xmark & flag      & \xmark \\
PentestGPT~\cite{pentestgpt24}       & 1 host  & \xmark & task-step & \xmark \\
VulnBot~\cite{vulnbot25}             & 1 host  & \xmark & task-step & \xmark \\
AutoPentester~\cite{autopentester25} & 1 host  & \xmark & task-step & \xmark \\
CAI~\cite{cai25}                     & 1 host  & \xmark & flag      & \xmark \\
Pentest-R1~\cite{pentestr125}        & 1 host  & \xmark & task-step & \xmark \\
xOffense~\cite{xoffense25}           & 1 host  & \xmark & task-step & \xmark \\
EnIGMA~\cite{enigma25}               & 1 host  & \xmark & flag      & \xmark \\
CheckMate~\cite{checkmate25}         & 1 host  & \xmark & task-step & \xmark \\
\midrule
\multicolumn{5}{@{}l}{\emph{Evaluation corpora}} \\
\midrule
AutoPenBench~\cite{autopenbench24}   & 1 host  & \xmark & flag      & \xmark \\
CAIBench / RCTF2~\cite{caibench25}   & 1 host  & partial & flag     & \xmark \\
Vulhub~\cite{vulhub}                 & 1 host  & \xmark & flag      & \xmark \\
Ludus + GOAD~\cite{ludus,goad}       & network & \xmark & none      & \xmark \\
\midrule
\textbf{\systemname\ (ours)} & \textbf{network} & \textbf{\cmark} & \textbf{per-vuln} & \textbf{\cmark} \\
\bottomrule
\end{tabular}
\end{table}

% tab:reported removed — headline numbers from prior work are discussed
% in §2.1 prose; cross-benchmark comparison is Table~\ref{tab:cross} in §7.

% ══════════════════════════════════════════════════════════════════════════
% §3 Threat Model
% ══════════════════════════════════════════════════════════════════════════
\section{Threat Model and Problem Definition}
\label{sec:threat}

\medskip\noindent\textbf{Problem statement.}
Given a target subnet with a designated entry point~$e$ declared in each scenario's topology metadata, \emph{network-scale penetration testing} is the task of discovering all hosts reachable from~$e$, identifying their vulnerabilities, producing evidence of exploitability, and reporting findings across the full network. The unit of evaluation is the network, not an individual host: IoT risk is driven by pivot chains (e.g., entry point $\to$ gateway $\to$ MQTT broker $\to$ OT controller), and single-host metrics are blind to vulnerabilities that only become reachable after one or more pivots.

\medskip\noindent\textbf{Attacker model.}
The attacker is an external adversary with Layer-3 access to the target subnet but no prior credentials, no out-of-band knowledge of deployed software, and no privileged vantage beyond active scanning. Their objective is data exfiltration (sensor readings, credentials) or OT disruption (reaching industrial-control services behind a segmentation boundary). They may use standard pentest tooling (Nmap, MQTT clients, SSH, HTTP) augmented by an LLM with tool-calling, within a bounded turn budget (\S\ref{sec:harness:cost}). This adversarial setting requires multi-step reasoning across heterogeneous protocols, which motivates the use of an LLM-driven agent rather than a static scanner.

\medskip\noindent\textbf{Evidence levels.}
A finding is solved at the deepest level the agent can demonstrate 
within its turn budget:
\begin{itemize}
  \item \textsf{L1}~\emph{detection}: a port scan or banner confirms 
  the issue (e.g., unauthenticated MQTT port \texttt{1883} open).
  \item \textsf{L2}~\emph{exploitation}: an active interaction 
  demonstrates it (e.g., anonymous MQTT subscribe on \texttt{\#}, 
  SSH login with default credentials).
  \item \textsf{L3}~\emph{exfiltration}: sensitive content is retrieved
  (e.g., MariaDB \texttt{root} dump, camera RTSP stream).
\end{itemize}
This ordering governs both \textsc{LANCE} Phase~4 prompting (\S\ref{sec:harness:pipeline}) and evaluator scoring (\S\ref{sec:iotbench:eval}).

\medskip\noindent\textbf{Scope.}
We evaluate at the IP/application layer on virtual IoT networks. Out of scope: physical and side-channel attacks, RF-layer exploitation, firmware extraction, post-exploitation persistence, and insider threats.

% ══════════════════════════════════════════════════════════════════════════
% §4 IoTBench
% ══════════════════════════════════════════════════════════════════════════
\section{\systemname: A Network-Scale Benchmark}
\label{sec:iotbench}

\subsection{Design Principles}
\label{sec:iotbench:principles}

\textsc{IoTChainBench} is built around six principles:

\begin{itemize}
  \item \textbf{Reproducibility.} Full infrastructure deploys from Ansible playbooks against a Proxmox cluster; vulnerabilities are injected idempotently with no manual steps.

  \item \textbf{Architectural diversity.} Scenarios span five network patterns with reachability enforced by OpenWrt firewall rules in segmented/VLAN cases.

  \item \textbf{Protocol coverage.} Each scenario mixes IoT and OT protocols (MQTT, Modbus, LoRaWAN, CoAP, SNMP, SSH, HTTP, FTP, Telnet) to exercise protocol-specific reasoning.

  \item \textbf{Ground-truth granularity.} Every vulnerability is documented in YAML with device, IP, severity, OWASP IoT and MITRE ICS mappings, optional CVE, indicators, and a verification command.

  \item \textbf{Scalability axis.} Ten main scenarios (4--8 devices) and three scalability scenarios (15--35 devices) separate pattern coverage from network size.

  \item \textbf{Hardened controls.} Two vulnerability-free topologies reusing \texttt{s\_1} and \texttt{s\_4} measure agent False Positive rate.
\end{itemize}


\subsection{Network Topologies and Scenario Catalogue}
\label{sec:iotbench:patterns}

\systemname{} defines five topological patterns that vary the number of pivots required to reach the deepest vulnerable service, directly setting the MHR$_k$ difficulty axis:

\begin{enumerate}[nosep]
  \item \textbf{P1 — Flat.} Single subnet; all devices directly reachable from the entry point.
  \item \textbf{P2 — Gateway-centric.} One IoT gateway mediates all back-end access; back-end services unreachable without gateway compromise.
  \item \textbf{P3 — Segmented IT/OT.} Admin/IT/OT zones with ACLs; the OT zone (Modbus PLCs, HMI) requires a jump-host pivot.
  \item \textbf{P4 — Hub-centric star.} A central hub is the sole path to the MQTT broker, DB, and camera; peripherals are mutually isolated.
  \item \textbf{P5 — Edge-cloud pivot.} Edge and cloud subnets are separated; only the edge gateway bridges them.
\end{enumerate}

The 15 scenarios (Table~\ref{tab:scenarios}) are organised along three axes: 10~\emph{main} scenarios instantiate the patterns above at 4--8 devices; 3~\emph{scalability} stressors push to 15, 17, and 35 devices using logical multi-zone and real-VLAN variants; and 2~\emph{hardened controls} replicate the topologies of s\_1 and s\_4 with no injected vulnerability, used to measure agent false-positive rate.

\begin{table}[h!]
\centering
\caption{The 15 \systemname\ scenarios: 13 vulnerability-injected
(133 devices, 229 vulnerabilities), plus 2 hardened controls
(12 devices, 0 vulnerabilities).}
\label{tab:scenarios}
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}llcrl@{}}
\toprule
ID & Pattern / role & Dev. & Vulns & Theme \\
\midrule
\multicolumn{5}{@{}l}{\emph{Main scenarios (pattern coverage)}} \\
\midrule
\texttt{s\_1}  & P1 Flat              & 4  & 12 & Minimal IoT (MQTT/Web/SSH) \\
\texttt{s\_2}  & P1 Flat              & 6  & 13 & Flat IoT + IT mix \\
\texttt{s\_3}  & P2 Gateway-centric   & 8  & 18 & NATO lab replica \\
\texttt{s\_4}  & P3 Segmented IT/OT   & 8  & 18 & Industrial (Modbus/HMI) \\
\texttt{s\_5}  & P4 Hub-centric star  & 8  & 15 & Smart building (cameras/NVR) \\
\texttt{s\_6}  & P4 Hub-centric star  & 6  & 16 & Home automation hub \\
\texttt{s\_7}  & P5 Edge-cloud pivot  & 6  & 14 & Edge $\to$ cloud API \\
\texttt{s\_8}  & P3 Segmented IT/OT   & 8  & 14 & IT/IoT/OT 3-zone industrial \\
\texttt{s\_9}  & P1 Flat (mesh)       & 6  & 11 & Mesh IoT smart-city outdoor \\
\texttt{s\_10} & P1 Flat (variants)   & 6  & 13 & Role-variant stress \\
\midrule
\multicolumn{5}{@{}l}{\emph{Scalability stressors}} \\
\midrule
\texttt{s\_11} & Smart City (flat)    & 15 & 23 & 3 zones same /24 (no isolation) \\
\texttt{s\_13} & VLAN-segmented       & 17 & 20 & Real Proxmox VLAN bridge (3 VLANs) \\
\texttt{s\_12} & Smart City (large)   & 35 & 42 & Largest network (34 LXC services) \\
\midrule
\textbf{Total} &                      &\textbf{133} & \textbf{229} & \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Deterministic Vulnerability Injection}
\label{sec:iotbench:inject}

All vulnerabilities are injected by a dedicated Ansible role parameterized by \texttt{scenario\_id}, ensuring fully reproducible and idempotent deployments. Each scenario YAML declares router-level weaknesses (Telnet, WAN admin, FTP) and per-service role definitions that trigger targeted injections, covering the most common real-world IoT misconfigurations: anonymous MQTT access, weak SSH credentials, and passwordless database roots. Every injected vulnerability is mapped to OWASP IoT Top~10 and MITRE ATT\&CK for ICS, providing a structured taxonomy for evaluation.

The 229 injected vulnerabilities span 11 categories (Table~\ref{tab:categories}), dominated by misconfigurations (68), missing authentication (41), and data exposure (38). Severity skews high: 107 HIGH and 37 CRITICAL instances, reflecting realistic IoT deployment weaknesses rather than theoretical edge cases.

\begin{table}[h!]
\centering
\caption{Vulnerability category and severity distribution
(229 instances across 13 scenarios).}
\label{tab:categories}
\footnotesize
\begin{tabular}{@{}lrrrrr@{}}
\toprule
Category & Count & CRIT & HIGH & MED & LOW \\
\midrule
\texttt{misconfiguration}      & 68  &  7 & 36 & 25 &  0 \\
\texttt{no\_authentication}    & 41  & 17 & 21 &  3 &  0 \\
\texttt{data\_exposure}        & 38  &  0 & 14 & 24 &  0 \\
\texttt{default\_credentials}  & 29  &  3 & 26 &  0 &  0 \\
\texttt{info\_disclosure}      & 13  &  0 &  0 &  0 & 13 \\
\texttt{code\_injection}       & 11  & 10 &  1 &  0 &  0 \\
\texttt{cve}                   &  9  &  0 &  3 &  6 &  0 \\
\texttt{insecure\_update}      &  9  &  0 &  4 &  5 &  0 \\
\texttt{weak\_crypto}          &  6  &  0 &  0 &  1 &  5 \\
\texttt{missing\_header}       &  3  &  0 &  0 &  0 &  3 \\
\texttt{privilege\_escalation} &  2  &  0 &  2 &  0 &  0 \\
\midrule
\textbf{Total} & \textbf{229} & 37 & 107 & 64 & 21 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Ground-Truth Schema}
\label{sec:iotbench:gt}

Each scenario ships a YAML file with one entry per injected vulnerability. Each entry records the affected device, IP, role, severity, category, OWASP IoT Top~10 and MITRE ATT\&CK for ICS mappings, an optional CVE, expected detection indicators, and a verification command that confirms exploitability. This per-vulnerability granularity, absent from flag-capture benchmarks, enables the evidence-level and MHR$_k$ metrics defined in \S\ref{sec:iotbench:eval}.

\begin{figure}[h!]
\footnotesize
\begin{verbatim}
- id: V1
  title: "MQTT without authentication"
  severity: high
  category: misconfiguration
  owasp_iot: "I1 / I9"
  mitre_ics: "Initial Access, Collection"
  indicators:
    - "allow_anonymous true"
  verification: "mosquitto_sub -h 192.168.100.11 -t '#' -v"
\end{verbatim}
\caption{Ground-truth entry example (fields trimmed for brevity).}
\label{lst:gt}
\end{figure}

Scenarios may additionally declare a \texttt{bonus\_types} list of vulnerability types adjacent to the injected set. Agent findings that match these types are counted as neither True Positives nor False Positives, avoiding penalties for legitimate hardening recommendations while keeping the recall denominator fixed.

\subsection{Scoring and Evaluation Metrics}
\label{sec:iotbench:eval}

\medskip\noindent\textbf{Finding matcher.}
Each LLM finding is matched against ground truth in three ordered passes: (1)~exact CVE match; (2)~\texttt{(ip, vuln\_type)} with canonicalised type; (3)~\texttt{(ip, severity)} fallback. Unmatched findings are False Positives. Duplicate findings on \texttt{(ip, type, port)} are resolved by retaining the lower severity to counteract inflation observed in pilot runs.

\medskip\noindent\textbf{Weighted score.}
The composite score is a severity-weighted True-Positive sum ($w$: critical~4, high~3, medium~2, low~1) normalized by $\sum_{g \in \text{GT}} w(g)$, with $0.5\times$ and $0.75\times$ penalties for fallback and severity-mismatch matches respectively.

\medskip\noindent\textbf{Evidence levels.}
Per True Positive, the evaluator tracks the deepest level achieved: \textsf{L1}~detection, \textsf{L2}~exploitation, \textsf{L3}~exfiltration. $$\text{L}k\text{-coverage} = \frac{|\{t \in \text{TP} : \text{level}(t) \geq k\}|}{|\text{TP}|}$$ This metric is absent from binary flag-capture benchmarks.

\medskip\noindent\textbf{Multi-Hop Reach.}
For each ground-truth vulnerability $g$, hop depth counts the minimum number of pivots from entry point $e$: $$\text{hop\_depth}(g) = \min_{p \,\in\, \text{paths}(e \rightarrow d_g)} \max(0,\,|p|_E - 1)$$ Multi-Hop Reach at depth $k$ restricts recall to deep findings: $$\text{MHR}_k = \frac{|\{g \in \text{GT} : \text{hop\_depth}(g) \geq k \,\wedge\, g \in \text{TP}\}|}{|\{g \in \text{GT} : \text{hop\_depth}(g) \geq k\}|}$$ MHR$_1$, MHR$_2$, MHR$_3$ measure recall at increasing pivot depth. Agents limited to directly exposed hosts score zero for $k \geq 1$.

% ══════════════════════════════════════════════════════════════════════════
% §5 Harness
% ══════════════════════════════════════════════════════════════════════════
\section{\harnessname: A Network-Native LLM Pentest Harness}
\label{sec:harness}

Existing LLM pentest harnesses treat the network as a flat list of targets, letting a single agent accumulate unbounded context as the network grows. \textsc{LANCE} instead decomposes network pentesting into six phases with typed deliverables. The key design decision is \emph{parallelization at the device and vulnerability granularity}: rather than a single monolithic agent holding the full network state, we spawn dedicated micro-agents per target with constrained tool sets and aggregate deterministically. This keeps each micro-agent's context bounded at $O(1)$ in the number of network devices, with cross-device state re-introduced only in Phase~5 (intrusion) and Phase~6 (report). Vulnerability type strings produced by LLM agents are resolved through a canonical taxonomy of 11~categories (Table~\ref{tab:categories}), filtering noise and deduplicating collisions before evaluation.

\subsection{Pipeline}
\label{sec:harness:pipeline}

\noindent\textbf{Phase 1: Topology prior.} Phase~1 loads a directed graph $G=(V,E)$ of the target network, where $V$ are devices and $E$ are protocol-labeled links. In \emph{scenario mode}, the topology is loaded from the benchmark YAML; in \emph{discovery mode}, the graph starts empty and is populated by Phase~2. The output enumerates the attack surface and ranks pivot candidates by betweenness centrality (identifying devices whose compromise enables the most lateral paths).

\medskip\noindent\textbf{Phase 2: Network discovery.}
Phase~2 executes active reconnaissance across the target \texttt{/24} using host discovery, service/version scanning, L2 vendor probing, and protocol-specific probes (MQTT, SSH, HTTP). Unlike single-host harnesses, the agent derives the target IP from the CIDR scope rather than receiving it directly.

\medskip\noindent\textbf{Phase 3: Per-device parallel triage.} 
Phase~3 runs a deterministic scanner pass per device, then spawns one LLM micro-agent per device with only that device's scan output and a minimal tool set (\texttt{http\_get}, \texttt{cve\_search}). Per-device outputs are merged and deduplicated via severity-aware rules. This decomposition keeps each agent's context constant regardless of network size, at the cost of cross-device correlation, which Phase~5 partially recovers through lateral movement.

\medskip\noindent\textbf{Phase 4: Per-vulnerability verification.}
Phase~4 spawns one micro-agent per Phase~3 finding with full operational tool access (\texttt{ssh\_login}, \texttt{mysql\_query}, \texttt{mqtt\_listen}, \texttt{modbus\_scan}, \texttt{telnet\_connect}). Each agent attempts the deepest evidence level achievable (\textsf{L1}--\textsf{L3}, \cref{sec:threat}) within a bounded turn budget. Phase~3 confirmed findings override Phase~4 errors to avoid erasing directly-observed indicators.

\medskip\noindent\textbf{Phase 5: Intrusion and lateral movement.}
Phase~5 turns the per-device vulnerability inventory into a network-scoped infiltration campaign. The intrusion agent attempts credential spraying, lateral movement along edges present in the topology graph from Phase~1 (reflecting observed or asserted reachability), and crown-jewel access on back-end services unreachable from the entry point. Every hop is recorded with source, destination, protocol, and credentials used. This phase is the sole producer of \textsf{L3} evidence and MHR$_k$ findings at $k \geq 2$; ablating it forces those columns to zero by construction.

\medskip\noindent\textbf{Phase 6: Network report.}
Phase~6 produces a network-scoped report with device inventory, per-device vulnerabilities, attack surface summary, multi-hop intrusion narrative, and recommendations ordered by severity $\times$ exposure. Cross-device aggregation is deliberately deferred to this phase to avoid bloating earlier agents, and the consolidated output enables a human analyst to prioritize remediation across the full network rather than per-host.

\subsection{Cost Envelope}
\label{sec:harness:cost}

For a network with $n$ devices and $v$ findings per device on average, the harness issues $O(nv)$ LLM invocations:
$$\underbrace{1}_{P1} + \underbrace{1}_{P2} + \underbrace{n}_{P3} + \underbrace{nv}_{P4} + \underbrace{1}_{P5} + \underbrace{1}_{P6}$$
Each invocation is bounded by a fixed turn budget (50 for P1--P2, 10 per device for P3, 10 per finding for P4, 30 for P5, 20 for P6). Cost scales linearly with network size, not combinatorially.

The three structural choices above, topology prior (P1), per-device parallelism (P3--P4), and the dedicated intrusion phase (P5), each correspond to an ablation evaluated in \cref{sec:eval}.

% ══════════════════════════════════════════════════════════════════════════
% §6 Experimental Setup
% ══════════════════════════════════════════════════════════════════════════
\section{Experimental Setup}
\label{sec:setup}

\medskip\noindent\textbf{Infrastructure.}
All 15 scenarios (145 devices total) run on a single Proxmox VE host, deployed as LXC containers on a dedicated Linux bridge with an OpenWrt router as gateway. Per-scenario deployment and teardown are fully automated via Ansible.

\medskip\noindent\textbf{Model.}
To isolate architectural contributions from model-specific advantages, all LLM-driven systems run on the same general-purpose model: \textbf{MiniMax-M2.7}~\cite{minimax27}. Security-fine-tuned models are intentionally excluded; this study addresses the architecture axis, not the model fine-tuning axis.

\medskip\noindent\textbf{Baselines.}
We compare \harnessname{} against two ablations isolating its key design choices, and one non-LLM scanner baseline:
\begin{itemize}
  \item \textbf{\harnessname{}--w/o-P1} disables the topology prior, testing the contribution of infrastructure-graph guidance.
  \item \textbf{\harnessname{}--w/o-P5} disables the intrusion phase, testing the contribution of lateral movement to \textsf{L3} evidence and MHR$_k$. Ablating P5 forces \textsf{L3}-coverage and MHR$_k \geq 2$ to zero by construction.
  \item \textbf{Nmap NSE} runs Nmap with all NSE scripts against the same target CIDR, converted to the evaluator's JSON schema.
\end{itemize}

\medskip\noindent\textbf{Metrics.}
We report five families: task completion (recall), quality (precision, F1, CVSS-weighted score), Multi-Hop Reach (MHR$_k$ for $k \in \{1,2,3\}$), evidence depth (\textsf{L1}/\textsf{L2}/\textsf{L3} coverage), and cost (tokens, USD, wall-clock time); the first four are formally defined in \cref{sec:iotbench:eval}. The last three are absent from existing single-host benchmarks and directly measure the network-scale capabilities \harnessname{} is designed to exercise.

\medskip\noindent\textbf{Variance.}
The evaluator supports repeated runs per (system, scenario) cell with mean, standard deviation, bootstrap confidence intervals, and pairwise Mann--Whitney $U$ tests on F1 and weighted score. We report scenario-level means; confidence intervals are included where multiple runs per cell are available.

% ══════════════════════════════════════════════════════════════════════════
% §7 Evaluation
% ══════════════════════════════════════════════════════════════════════════
\section{Evaluation}
\label{sec:eval}

\subsection{Overall performance}
\label{sec:eval:overall}

\begin{table*}[t]
\centering
\caption{Results on \systemname\ (10 main scenarios, MiniMax-M2.7). TC = Recall;
Score = CVSS-weighted Recall. Baseline rows report the F1 values available in
the scenario comparison export; the remaining columns are not reported by those
adapters.}
\label{tab:main}
\footnotesize
\setlength{\tabcolsep}{6pt}
\begin{tabular}{@{}lrrrrrl@{}}
\toprule
System & TC & Prec. & F1 & Score & Wins & Notes \\
\midrule
\textbf{\harnessname (full)} & \textbf{0.962} & \textbf{0.922} & \textbf{0.940} & \textbf{86.1\%} & \textbf{10/10} & full pipeline \\
\midrule
CAI-style adapter      & -- & -- & 0.312 & -- & 0/10 & same scenario export \\
PentGPT-style adapter  & -- & -- & 0.321 & -- & 0/10 & same scenario export \\
VulnBot-style adapter  & -- & -- & 0.316 & -- & 0/10 & same scenario export \\
Best baseline per scenario & -- & -- & 0.393 & -- & 0/10 & oracle best of the three \\
\bottomrule
\end{tabular}

\smallskip
\footnotesize\emph{Means over scenarios \texttt{s\_1}--\texttt{s\_10}. Wins count
per-scenario F1 wins over \harnessname{} for baseline rows, and
\harnessname{} wins over the best baseline for the first row.}
\end{table*}

Across the 10 main scenarios of \systemname{}, \harnessname{} achieves a recall
of 0.962, precision of 0.922, and F1 of 0.940, with a CVSS-weighted score of
86.1\% (Table~\ref{tab:main}). The result is not driven by a single easy topology:
\harnessname{} outperforms the strongest baseline in every individual scenario,
with a mean absolute F1 gap of 0.547 over the per-scenario oracle best baseline.
The two additional scalability scenarios (\texttt{s\_11}, \texttt{s\_12}) are
consistent with this trend, reaching a mean F1 of 0.909 and a mean weighted
score of 87.7\%.

\begin{figure}[h!]
  \centering
  \begin{tikzpicture}[
    font=\scriptsize,
    ours/.style={draw=blue!80!black, line width=0.4pt, fill=blue!60},
    base/.style={draw=gray!60,       line width=0.4pt, fill=gray!35},
    best/.style={draw=gray!60,       line width=0.4pt, fill=gray!35, dashed},
    grid/.style={draw=black!12, line width=0.2pt},
    axis/.style={draw=black!55, line width=0.4pt}
  ]
    \draw[axis] (0,0) -- (9.0,0);
    \draw[axis] (0,0) -- (0,3.85);
    \foreach \y/\lab in {0/0, 0.875/25, 1.75/50, 2.625/75, 3.5/100} {
      \draw[grid] (0,\y) -- (9.0,\y);
      \node[anchor=east,font=\tiny] at (-0.1,\y) {\lab};
    }
    \node[rotate=90,anchor=south,font=\tiny] at (-0.55,1.75) {Mean F1 (\%)};
    % LANCE
    \draw[ours] (0.25,0) rectangle +(0.9,3.290);
    \node[anchor=south,font=\scriptsize\bfseries,text=blue!70!black] at (0.70,3.33) {94.0};
    \node[anchor=north,font=\scriptsize] at (0.70,-0.14) {\harnessname{}};
    % CAI
    \draw[base] (2.0,0) rectangle +(0.9,1.092);
    \node[anchor=south,font=\scriptsize] at (2.45,1.13) {31.2};
    \node[anchor=north,font=\scriptsize] at (2.45,-0.14) {CAI};
    % PentGPT
    \draw[base] (3.75,0) rectangle +(0.9,1.124);
    \node[anchor=south,font=\scriptsize] at (4.20,1.16) {32.1};
    \node[anchor=north,font=\scriptsize] at (4.20,-0.14) {PentGPT};
    % VulnBot
    \draw[base] (5.5,0) rectangle +(0.9,1.106);
    \node[anchor=south,font=\scriptsize] at (5.95,1.14) {31.6};
    \node[anchor=north,font=\scriptsize] at (5.95,-0.14) {VulnBot};
    % Best baseline (oracle)
    \draw[best] (7.25,0) rectangle +(0.9,1.376);
    \node[anchor=south,font=\scriptsize] at (7.70,1.41) {39.3};
    \node[anchor=north,font=\scriptsize,align=center] at (7.70,-0.14) {Best\\baseline};
  \end{tikzpicture}
  \caption{Mean F1 on the 10 main \systemname{} scenarios (MiniMax-M2.7).
  \harnessname{} (blue) outperforms each adapted LLM-pentest baseline and the
  per-scenario oracle best (dashed). The gap of 0.547 F1 points is consistent
  across all 10 scenarios.}
  \label{fig:baselines}
\end{figure}

Figure~\ref{fig:baselines} summarises the comparison graphically. The three adapted
baselines (CAI-style, PentGPT-style, VulnBot-style) each score below 0.322 F1
on average, and their per-scenario best reaches only 0.393. The gap of 0.547
F1 points reflects a structural advantage rather than a single difficult
scenario: baselines succeed only on those scenarios where the vulnerability set
overlaps with patterns they probe by default, while \harnessname{}'s
topology-aware pipeline and deterministic scanner cover the full injected set.

\subsection{Cross-benchmark comparison on single-host corpora}
\label{sec:eval:baselines}

\begin{table*}[t]
\centering
\caption{Cross-benchmark comparison. Published single-host numbers are shown
for reference; \harnessname{} (MiniMax-M2.7) is evaluated on both AutoPenBench
and Vulhub. Models differ across rows: this table is not a fair model
comparison but checks whether \harnessname{} remains competitive on established
single-host corpora while adding the network-scale dimension.
\emph{E2E \%}: fraction of tasks fully solved. CheckMate uses milestone-based
progress on a 120-container Vulhub sample and reports only a relative
success-rate gain over its baseline. ``n/r'' = not reported.
$^\star$ = lower-bound proof-oracle estimate (see text).}
\label{tab:cross}
\setlength{\tabcolsep}{6pt}
\footnotesize
\begin{tabular}{@{}lllrr@{}}
\toprule
System & Model & Corpus & Sub-task \% & E2E \% \\
\midrule
\multicolumn{5}{@{}l}{\emph{AutoPenBench (33 tasks)}} \\
\midrule
Llama 3.1-405B baseline~\cite{vulnbot25}  & Llama 3.1-405B    & AutoPenBench & 49.1 & 9.1  \\
GPT-4o baseline~\cite{vulnbot25}          & GPT-4o            & AutoPenBench & n/r  & 21.2 \\
\textbf{VulnBot}~\cite{vulnbot25}         & Llama 3.1-405B    & AutoPenBench & \textbf{69.1} & \textbf{30.3} \\
\textbf{xOffense}~\cite{xoffense25}       & Qwen3-32B (LoRA)  & AutoPenBench & \textbf{79.2} & n/r  \\
\textbf{\harnessname\ (ours, blind)}      & MiniMax-M2.7      & AutoPenBench & 39.4$^\dagger$ & 21.2 \\
\textbf{\harnessname\ (ours, informed)}   & MiniMax-M2.7      & AutoPenBench & 51.5$^\dagger$ & 27.3$^\ddagger$ \\
\midrule
\multicolumn{5}{@{}l}{\emph{Vulhub / Vulhub-derived single-host tasks}} \\
\midrule
Claude Code (baseline)~\cite{checkmate25} & Sonnet 4.5        & Vulhub & n/r & n/r \\
\textbf{CheckMate}~\cite{checkmate25}     & Sonnet 4.5 + Classical Planning & Vulhub & n/r & \textbf{+20\%\textsuperscript{rel}} \\
\textbf{\harnessname\ (ours)}             & MiniMax-M2.7      & Vulhub & n/r & \textbf{39--48\%}$^\star$ \\

\bottomrule
\end{tabular}
\smallskip\\
\footnotesize\emph{%
\textsuperscript{rel} Relative benchmark success-rate gain over Claude Code
baseline; absolute E2E rate not published by CheckMate.
$^\dagger$ We report useful findings as the closest sub-task analogue because
our harness does not use AutoPenBench's upstream orchestrator. $^\ddagger$
Informed mode receives the benchmark case id and vulnerability label; blind
mode is the fair no-hint setting.
$^\star$ Full 191-case distributed run. The automatic useful-finding lower
bound is 69/175 completed agent runs (39\%); manual inspection identifies 15
additional self-reported \textsc{no finding} cases with exploit evidence,
raising the conservative lower bound to 84/175 (48\%).}
\end{table*}

\medskip\noindent\textbf{Vulhub.}
To assess transfer beyond IoT network scenarios, we evaluate \harnessname{} on
the Vulhub single-host corpus. The distributed run processed all 191 cases
across six workers. Of these, 175 reached a completed agent run; 11 failed
during environment setup; 3 were interrupted by provider rate limits; and 2 were
skipped after the rate-limit breaker tripped. The 175 completed runs consumed
194.5M tokens at an estimated cost of \$60.96, or approximately \$0.33 per
executed case.

The automatic verdict labels 69 of the 175 completed cases as useful findings
(39\%). This figure undercounts success because Vulhub does not provide a
uniform flag oracle: an agent that demonstrates command execution or file
disclosure may still self-report failure while continuing to search for a
nonexistent flag file. Re-reading the \texttt{proof.json} outputs identifies
15 such cases, raising the conservative lower bound to 84/175 (48\%). The six
cases with strictly confirmed exploitation are \texttt{struts2/s2-053},
\texttt{phpunit/CVE-2017-9841}, \texttt{redis/CVE-2022-0543},
\texttt{jenkins/CVE-2018-1000861}, \texttt{php/CVE-2024-2961}, and
\texttt{spring/CVE-2022-22947}; the remaining useful cases are either probable
vulnerabilities or tool-blocked and credential-blocked findings with actionable
evidence.

\medskip\noindent\textbf{AutoPenBench.}
The AutoPenBench rows use the same MiniMax-M2.7 model and the same \harnessname{}
pipeline, but differ in context policy. In \emph{blind} mode, the agent receives
only the target endpoint and exposed service metadata. It captures 7/33 flags
(21.2\%) and produces useful findings on 13/33 tasks (39.4\%), consuming 15.5M
tokens at a cost of \$4.88. In \emph{informed} mode, the agent additionally
receives the benchmark case identifier and vulnerability label, raising flag
capture to 9/33 (27.3\%) and useful findings to 17/33 tasks (51.5\%) at a cost
of \$7.43 and 23.6M tokens. Blind mode is the fair external evaluation; informed
mode provides an upper-bound diagnostic on the value of vulnerability-label
hints. In both configurations, two real-world CVE tasks fail before agent
invocation due to upstream environment issues: one involves a removed Docker
base image and one involves a container that enters a crash loop before the
agent can connect.

Compared with published baselines on AutoPenBench, \harnessname{} in blind mode
matches the GPT-4o single-agent baseline at 21.2\% E2E, despite using a
different model and a multi-agent architecture not tuned for single-host tasks.
The gap relative to VulnBot (30.3\%) and xOffense (79.2\% sub-task) is
consistent with \harnessname{} being designed and evaluated primarily for
network-scale scenarios: its tool set and phase structure target multi-hop
reasoning rather than single-host exploit chaining.

\subsection{Per-scenario analysis}
\label{sec:eval:perscenario}

\begin{figure}[h!]
  \centering
  \begin{tikzpicture}[font=\tiny]
    \def\cellw{0.60}
    \def\cellh{0.42}
    % column headers
    \foreach \c/\sid in {0/S1,1/S2,2/S3,3/S4,4/S5,5/S6,6/S7,7/S8,8/S9,9/S10} {
      \node[rotate=40, anchor=west, font=\scriptsize] at (\c*\cellw+0.08,0.32) {\sid};
    }
    % row labels
    \foreach \r/\sys in {0/\textbf{LANCE},1/CAI,2/PentGPT,3/VulnBot} {
      \node[anchor=east, font=\scriptsize] at (-0.08,-\r*\cellh-\cellh/2) {\sys};
    }
    % separator line between LANCE and baselines
    \draw[black!40, line width=0.6pt] (0,-\cellh) -- (10*\cellw,-\cellh);
    % LANCE row (blue scale, white text)
    \foreach \c/\v in {0/92,1/92,2/95,3/94,4/100,5/100,6/97,7/93,8/95,9/82} {
      \draw[fill=blue!\v!20!white, draw=black!20, line width=0.2pt]
        (\c*\cellw,0) rectangle ++(\cellw,-\cellh);
      \node[text=white, font=\bfseries\tiny] at (\c*\cellw+\cellw/2,-\cellh/2) {\v};
    }
    % CAI row (orange scale, floor at 8 for visibility)
    \foreach \c/\v/\cv in
      {0/15/15,1/35/35,2/0/8,3/36/36,4/50/50,5/40/40,6/50/50,7/55/55,8/17/17,9/14/14} {
      \draw[fill=orange!\cv!8!white, draw=black!20, line width=0.2pt]
        (\c*\cellw,-\cellh) rectangle ++(\cellw,-\cellh);
      \node[font=\tiny] at (\c*\cellw+\cellw/2,-\cellh-\cellh/2) {\v};
    }
    % PentGPT row
    \foreach \c/\v/\cv in
      {0/0/8,1/14/14,2/0/8,3/42/42,4/64/64,5/52/52,6/45/45,7/35/35,8/43/43,9/27/27} {
      \draw[fill=orange!\cv!8!white, draw=black!20, line width=0.2pt]
        (\c*\cellw,-2*\cellh) rectangle ++(\cellw,-\cellh);
      \node[font=\tiny] at (\c*\cellw+\cellw/2,-2*\cellh-\cellh/2) {\v};
    }
    % VulnBot row
    \foreach \c/\v/\cv in
      {0/15/15,1/45/45,2/0/8,3/44/44,4/50/50,5/40/40,6/45/45,7/35/35,8/29/29,9/14/14} {
      \draw[fill=orange!\cv!8!white, draw=black!20, line width=0.2pt]
        (\c*\cellw,-3*\cellh) rectangle ++(\cellw,-\cellh);
      \node[font=\tiny] at (\c*\cellw+\cellw/2,-3*\cellh-\cellh/2) {\v};
    }
    % outer border
    \draw[black!50, line width=0.5pt] (0,0) rectangle (10*\cellw,-4*\cellh);
    % legend
    \node[font=\tiny, anchor=east] at (10*\cellw+0.05,-4*\cellh-0.30) {F1 (\%):};
    \node[font=\tiny, anchor=west] at (10*\cellw+0.05,-4*\cellh-0.30) {0};
    \shade[left color=white, right color=blue!80, draw=black!30, line width=0.2pt]
      (10*\cellw+0.45,-4*\cellh-0.42) rectangle +(1.2,0.22);
    \node[font=\tiny, anchor=west] at (10*\cellw+1.70,-4*\cellh-0.30) {100};
  \end{tikzpicture}
  \caption{Per-scenario F1 heatmap on the 10 main \systemname{} scenarios (values
  in \%). \harnessname{} (blue, top row) achieves perfect F1 on hub-centric
  topologies (\texttt{S5}, \texttt{S6}) and is weakest on the flat role-variant
  scenario \texttt{S10}. Baselines (orange) are uniformly lower.}
  \label{fig:heatmap}
\end{figure}

The per-scenario breakdown in Figure~\ref{fig:heatmap} reveals a clear pattern. The
hub-centric star topologies (\texttt{s\_5}, \texttt{s\_6}) are not difficult:
both reach F1 = 1.000, which indicates that explicit graph context combined with
deterministic probing makes gateway-mediated dependencies tractable. Segmented
IT/OT scenarios remain strong but not perfect (\texttt{s\_4}: F1 = 0.944,
\texttt{s\_8}: F1 = 0.933), with residual errors caused by deeper pivots that
increase the chance of duplicate or slightly misclassified findings. The
edge-cloud pivot (\texttt{s\_7}) stays high at F1 = 0.965, and the mesh
smart-city case (\texttt{s\_9}) reaches F1 = 0.952 with no false positives.

The most difficult main scenario is \texttt{s\_10} (F1 = 0.815, Score = 71.3\%).
This topology is intentionally flat rather than multi-hop, so the degradation
is not caused by lateral-movement reasoning. Instead, service role variants
reduce the predictability of service semantics and increase the proportion of
near-duplicate reports. The two scalability extensions confirm stability:
\texttt{s\_11} achieves F1 = 0.875 and Score = 84.8\%, and \texttt{s\_12}
achieves F1 = 0.942 and Score = 90.6\%.

\subsection{Failure modes and cost}
\label{sec:eval:failures}

\begin{table}[t]
  \centering
\caption{Vulhub full-run outcome distribution. Percentages are over the 175
completed agent runs unless marked as pre-agent failures.}
\label{tab:vulhub-failures}
\scriptsize
\begin{tabular}{@{}p{0.30\linewidth}rrp{0.42\linewidth}@{}}
\toprule
Outcome & Cases & Share & Interpretation \\
\midrule
Max turns & 83 & 47\% & loops and exploited-but-no-flag cases \\
No finding & 23 & 13\% & includes 15 hidden proof-level successes \\
Blocked: credentials & 23 & 13\% & exploit path required unavailable secret \\
Probable vulnerability & 21 & 12\% & actionable but unconfirmed \\
Blocked: tool & 19 & 11\% & missing exploit or runtime support \\
Confirmed exploit & 6 & 3\% & strict automatic success \\
\midrule
Environment failed & 11 & -- & target did not reach the agent \\
Rate-limited / skipped & 5 & -- & provider-side interruption \\
\bottomrule
\end{tabular}
\end{table}

\medskip\noindent\textbf{Failure modes.}
On \systemname{} main scenarios, failures are accuracy failures rather than
resource failures. The \texttt{s\_10} scenario dominates the error profile,
contributing both missed findings and near-duplicate false positives that arise
from role variants. The remaining nine main scenarios each stay above F1 = 0.923,
suggesting that the pipeline generalises well across varied topologies.

On Vulhub (Table~\ref{tab:vulhub-failures}), the dominant outcome is
\textsc{max turns} (47\%), which conflates two behaviours: genuinely
unproductive search loops and successful exploitation followed by continued
searching for a flag that the Vulhub environment does not standardise. The next
most actionable failure categories are tool coverage (19 cases blocked by
missing exploit support) and credential handling (23 cases requiring secrets
the agent did not possess).

\medskip\noindent\textbf{Computational cost.}
The 10 main \systemname{} scenarios run at a cost that scales with network size.
On the AutoPenBench single-host corpus, blind-mode evaluation costs \$4.88 and
consumes 15.5M tokens for 33 tasks (\$0.15 per task); informed mode costs
\$7.43 and 23.6M tokens (\$0.23 per task). The Vulhub distributed run
(191 cases, six workers) consumed 194.5M tokens at a total cost of \$60.96,
approximately \$0.33 per executed case. These figures confirm that \harnessname{} remains within a practical per-run
budget for research deployments.

% ══════════════════════════════════════════════════════════════════════════
% §8 Discussion
% ══════════════════════════════════════════════════════════════════════════
\section{Discussion and Limitations}
\label{sec:discussion}

\medskip\noindent\textbf{Where LLMs help, where they don't.}
\systemname{} is designed to separate failure modes that binary flag-capture
benchmarks collapse into a single signal. The results show strong aggregate
vulnerability matching across varied topologies, with the per-scenario heatmap
confirming that the gain is not driven by a single easy pattern. The intended
deployment model is hybrid: LLM agents for configuration auditing and
multi-step reasoning, traditional scanners for broad CVE coverage, and human
operators or specialised tooling for the OT/ICS layer.

\medskip\noindent\textbf{Implications for agentic systems.}
The harness's per-device and per-vulnerability micro-agent structure is one
answer to context scaling, but it cedes cross-device inference. Designs that
keep a global aggregator agent in the loop without paying the full
network-context cost are an open research direction.

\medskip\noindent\textbf{Limitations.}
(i)~\emph{Simulated infrastructure.} \systemname{} scenarios emulate IoT
devices on Linux VMs; real constrained hardware introduces embedded-specific
attack surfaces (bootloaders, RTOS, firmware) we do not cover.
(ii)~\emph{Evaluator leniency.} Bonus categories are accepted as non-FP even
without ground-truth matches, which inflates precision.
(iii)~\emph{Model variance.} All results reported here are single-run estimates;
the evaluator supports repeated runs with bootstrap confidence intervals and
Mann--Whitney tests, but this study reports scenario-level means over one run
per (system, scenario) cell.
(iv)~\emph{Prompt sensitivity.} We fix a single prompt family across models;
per-model prompt tuning would likely raise scores.
(v)~\emph{Static architectures.} The five patterns are fixed; real deployments
evolve over time and include devices not represented here (RFID badge readers,
industrial PLCs, medical devices).

% ══════════════════════════════════════════════════════════════════════════
% §9 Ethical Considerations
% ══════════════════════════════════════════════════════════════════════════
\section{Ethical Considerations}
\label{sec:ethical}

\medskip\noindent\textbf{Isolation.}
All experiments run on an isolated Proxmox cluster with a dedicated bridge
(\texttt{vmbr1}) and no route to external networks; the OpenWrt gateway's
\texttt{WAN} interface is firewall-isolated. No live hosts, production
services, or third-party infrastructure are probed.

\medskip\noindent\textbf{Responsible disclosure.}
\systemname{} vulnerabilities are deterministically injected into our own
lab VMs via Ansible. No 0-days are introduced; all categories (misconfig,
weak credentials, Terrapin CVE-2023-48795, outdated Dropbear) correspond to
patterns already publicly documented. No responsible-disclosure process is
triggered.

\medskip\noindent\textbf{Misuse risk of released artifact.}
The benchmark artifact (Ansible playbooks, ground truth, evaluator) does not
embed exploitation payloads for unknown vulnerabilities, and the injected
weaknesses are common misconfigurations documented in OWASP~IoT~Top~10.
The harness uses only standard pentest tools already in widespread use.
We judge the net societal benefit (reproducible evaluation of LLM
pentest agents, and a concrete evaluation surface for future defenders)
to exceed the incremental misuse risk.

\medskip\noindent\textbf{IRB / human subjects.}
No human subjects are involved. No personal data is collected or processed.

\medskip\noindent\textbf{LLM costs and environmental footprint.}
We report token-estimated cost, actual API billing, and wall-clock time with
the experimental results. Energy usage is not claimed unless measured from
provider- or host-side telemetry.

% ══════════════════════════════════════════════════════════════════════════
% §10 Conclusion
% ══════════════════════════════════════════════════════════════════════════
\section{Conclusion}
\label{sec:conclusion}

We presented \systemname, a reproducible benchmark for evaluating LLM
penetration testing agents at \emph{network scale} on IoT infrastructure, and
\harnessname, a network-native multi-agent harness whose infrastructure-graph
guidance and dedicated lateral movement phase keep LLM context tractable while
explicitly modelling reachability. \harnessname{} achieves 0.940 mean F1 and 86.1\% weighted score on the
10 main scenarios, outperforming each adapted LLM-agent baseline in every
scenario. The benchmark additionally defines MHR$_k$, per-protocol recall,
and L1--L3 evidence coverage metrics that expose architectural differences
invisible to binary flag-capture corpora.

\medskip\noindent\textbf{Future work.}
(i)~\emph{Multi-model robustness:} extending the architectural comparison to
recent frontier and open-weight models to confirm model-independence of the
architectural advantage;
(ii)~\emph{Real hardware:} extending to embedded Zigbee, LoRaWAN, and
sub-GHz attack surfaces via SDR tooling;
(iii)~\emph{Defender's side:} using \systemname{} as a benchmark for
LLM-assisted IoT defense (detection, patching recommendation).

All code, scenarios, ground truth, and run logs are released at
\texttt{ANONYMIZED\_URL} under an open license.

% ── bibliography ──────────────────────────────────────────────────────────
\begin{thebibliography}{00}

\bibitem{happe23} A.~Happe and J.~Cito, ``Getting pwn'd by AI: Penetration testing with large language models,'' in \textit{Proc. ACM ESEC/FSE}, 2023.

\bibitem{hackingbuddygpt} A.~Happe \textit{et al.}, ``HackingBuddyGPT: An LLM-driven pentest agent,'' 2024. [Online]. Available: https://github.com/ipa-lab/hackingBuddyGPT

\bibitem{pentestgpt24} G.~Deng \textit{et al.}, ``PentestGPT: Evaluating and harnessing large language models for automated penetration testing,'' in \textit{Proc. 33rd USENIX Security Symp.}, 2024.

\bibitem{autoattacker24} J.~Xu \textit{et al.}, ``AutoAttacker: A large language model guided system to implement automatic cyber-attacks,'' \textit{arXiv:2403.01038}, 2024.

\bibitem{nebula24} O.~V.~Aguinaga, ``Nebula: An LLM-powered penetration testing assistant,'' \textit{arXiv:2412.00006}, 2024.

\bibitem{vulnbot25} H.~He \textit{et al.}, ``VulnBot: Autonomous penetration testing for a multi-agent collaborative framework,'' \textit{arXiv:2501.13411}, 2025.

\bibitem{autopentester25} Y.~Ginige \textit{et al.}, ``AutoPentester: An LLM agent-based framework for automated pentesting,'' \textit{arXiv:2510.05605}, 2025.

\bibitem{pentestagent25} X.~Shen \textit{et al.}, ``PentestAgent: Incorporating LLM agents to automated penetration testing,'' in \textit{Proc. 20th ACM Asia Conf. Computer and Communications Security (ASIA CCS)}, 2025. doi: 10.1145/3708821.3733882.

\bibitem{cai25} V.~Mayoral-Vilches \textit{et al.}, ``CAI: An open, bug bounty-ready cybersecurity AI,'' \textit{arXiv:2504.06017}, 2025.

\bibitem{pentestr125} Anonymous, ``Pentest-R1: Towards autonomous penetration testing reasoning optimized via two-stage reinforcement learning,'' \textit{arXiv:2508.07382}, 2025.

\bibitem{xoffense25} Anonymous, ``xOffense: An AI-driven autonomous penetration testing framework with offensive knowledge-enhanced LLMs and multi-agent systems,'' \textit{arXiv:2509.13021}, 2025.

\bibitem{enigma25} T.~Abramovich \textit{et al.}, ``EnIGMA: Interactive tools substantially assist LM agents in solving CTF challenges,'' in \textit{Proc. ICML}, 2025.

\bibitem{iris25} Z.~Li, S.~Dutta, and M.~Naik, ``IRIS: LLM-assisted static analysis for detecting security vulnerabilities,'' in \textit{Proc. NDSS}, 2025.

\bibitem{chap26} Anonymous, ``Context relay for long-running penetration-testing agents,'' in \textit{NDSS Workshop on LLM-Assisted Security and Trust Exploration (LAST-X)}, 2026.

\bibitem{awe26} Anonymous, ``AWE: A memory-augmented multi-agent framework for autonomous web penetration testing,'' in \textit{NDSS Workshop on LLM-Assisted Security and Trust Exploration (LAST-X)}, 2026.

\bibitem{checkmate25} L.~Wang, X.~Shi, Z.~Li, Y.~Jiang, S.~Tan, Y.~Jiang, J.~Cheng, W.~Chen, X.~Shen, Z.~Li, and Y.~Chen, ``Automated penetration testing with LLM agents and classical planning,'' \textit{arXiv:2512.11143}, 2025.

\bibitem{autopenbench24} Anonymous, ``Towards automated penetration testing: Introducing LLM benchmark, analysis, and improvements,'' \textit{arXiv:2410.17141}, 2024.

\bibitem{nyuctfbench24} M.~Shao \textit{et al.}, ``NYU CTF Bench: A scalable open-source benchmark dataset for evaluating LLMs in offensive security,'' in \textit{Proc. NeurIPS Datasets and Benchmarks Track}, 2024.

\bibitem{cybench24} A.~K.~Zhang \textit{et al.}, ``Cybench: A framework for evaluating cybersecurity capabilities and risk of language models,'' \textit{arXiv:2408.08926}, 2024.

\bibitem{cybereval24} M.~Bhatt \textit{et al.}, ``CyberSecEval~2: A wide-ranging cybersecurity evaluation suite for large language models,'' \textit{arXiv:2404.13161}, 2024.

\bibitem{caibench25} V.~Mayoral-Vilches \textit{et al.}, ``CAIBench: A meta-benchmark for evaluating cybersecurity AI agents,'' \textit{arXiv:2510.24317}, 2025.

\bibitem{ludus} Ludus Cloud, ``Ludus: Cyber range automation on Proxmox.'' [Online]. Available: https://ludus.cloud

\bibitem{goad} M3, ``GOAD: Game of Active Directory.'' [Online]. Available: https://github.com/Orange-Cyberdefense/GOAD

\bibitem{vulhub} Vulhub contributors, ``Vulhub: Pre-built vulnerable Docker environments for learning and reproducing CVEs.'' [Online]. Available: https://github.com/vulhub/vulhub

\bibitem{antonakakis17} M.~Antonakakis \textit{et al.}, ``Understanding the Mirai botnet,'' in \textit{Proc. 26th USENIX Security Symp.}, 2017.

\bibitem{verkada21} W.~Turton, ``Hackers breach thousands of security cameras, exposing Tesla, jails, hospitals,'' \textit{Bloomberg}, Mar.~2021.

\bibitem{langner11} R.~Langner, ``Stuxnet: Dissecting a cyberwarfare weapon,'' \textit{IEEE Security \& Privacy}, vol.~9, no.~3, pp.~49--51, 2011.

\bibitem{phillips98} C.~Phillips and L.~P.~Swiler, ``A graph-based system for network-vulnerability analysis,'' in \textit{Proc. New Security Paradigms Workshop}, 1998.

\bibitem{sheyner02} O.~Sheyner, J.~Haines, S.~Jha, R.~Lippmann, and J.~M.~Wing, ``Automated generation and analysis of attack graphs,'' in \textit{Proc. IEEE Symp. Security and Privacy}, 2002.

\bibitem{mulval05} X.~Ou, S.~Govindavajhala, and A.~W.~Appel, ``MulVAL: A logic-based network security analyzer,'' in \textit{Proc. USENIX Security Symp.}, 2005.

\bibitem{owasp-iot} OWASP Foundation, ``OWASP Internet of Things Top~10,'' 2024. [Online]. Available: https://owasp.org/www-project-internet-of-things-top-10/

\bibitem{mitre-ics} MITRE Corporation, ``MITRE ATT\&CK for ICS,'' 2024. [Online]. Available: https://attack.mitre.org/matrices/ics/

\bibitem{nist80082} K.~Stouffer \textit{et al.}, ``Guide to operational technology security,'' National Institute of Standards and Technology, NIST SP\,800-82 Rev.\,3, 2023.

\bibitem{iot23} S.~Garcia, A.~Parmisano, and M.~J.~Erquiaga, ``IoT-23: A labeled dataset with malicious and benign IoT network traffic,'' Stratosphere Laboratory, 2020. [Online]. Available: https://www.stratosphereips.org/datasets-iot23

\bibitem{toniot20} A.~Alsaedi, N.~Moustafa, Z.~Tari, A.~Mahmood, and A.~Anwar, ``TON\_IoT telemetry dataset: A new generation dataset of IoT and IIoT for data-driven intrusion detection systems,'' \textit{IEEE Access}, vol.~8, pp.~165130--165150, 2020.

\bibitem{edgeiiotset22} M.~A.~Ferrag, O.~Friha, D.~Hamouda, L.~Maglaras, and H.~Janicke, ``Edge-IIoTset: A new comprehensive realistic cyber security dataset of IoT and IIoT applications for centralized and federated learning,'' \textit{IEEE Access}, vol.~10, pp.~40281--40306, 2022.

\bibitem{costin14} A.~Costin, J.~Zaddach, A.~Francillon, and D.~Balzarotti, ``A large-scale analysis of the security of embedded firmwares,'' in \textit{Proc. USENIX Security Symp.}, 2014.

\bibitem{firmadyne16} D.~D.~Chen, M.~Egele, M.~Woo, and D.~Brumley, ``Towards automated dynamic analysis for Linux-based embedded firmware,'' in \textit{Proc. NDSS}, 2016.

\bibitem{firmae20} M.~Kim, D.~Kim, E.~Kim, S.~Kim, Y.~Jang, and Y.~Kim, ``FirmAE: Towards large-scale emulation of IoT firmware for dynamic analysis,'' in \textit{Proc. Annual Computer Security Applications Conf. (ACSAC)}, 2020.

\bibitem{iotgoat} OWASP, ``IoTGoat: A deliberately insecure IoT firmware,'' 2024. [Online]. Available: https://github.com/OWASP/IoTGoat

\bibitem{minimax27} MiniMax, ``MiniMax-M2.7: A general-purpose large language model,'' 2026. [Online]. Available: https://www.minimax.io/models/text/m27

\end{thebibliography}

\section*{LLM Usage Statement}
LLMs were used for editorial assistance while preparing this manuscript, and
all generated text was inspected by the authors to ensure accuracy,
originality, and consistency with the underlying artifacts. LLMs are also part
of the studied methodology: \harnessname{} uses tool-calling LLM agents for
network reconnaissance, vulnerability analysis, exploit verification,
intrusion attempts, and report generation, as described in
\cref{sec:harness}, \cref{sec:setup}, and \cref{sec:eval}.

\end{document}
